---
title: "175 final project"
author: "Junyue Wang"
output: html_document
---
```{r}
Sys.setlocale('LC_ALL','C')
library(survival)
rossi <- read.table(url("http://socserv.mcmaster.ca/jfox/Books/Companion/data/Rossi.txt"),header=TRUE,stringsAsFactors = FALSE)
rossi <- rossi[c(1:10)]

rossi$fin[rossi$fin == "no"]<- 0
rossi$fin[rossi$fin == "yes"]<- 1
rossi$race[rossi$race == "other"] <- 0
rossi$race[rossi$race == "black"] <- 1

rossi$wexp[rossi$wexp == "no"] <- 0
rossi$wexp[rossi$wexp == "yes"] <- 1
rossi$wexp <- as.numeric(rossi$wexp)

rossi$mar[rossi$mar == "not married"] <- 0
rossi$mar[rossi$mar == "married"] <- 1
rossi$mar <- as.numeric(rossi$mar)

rossi$paro[rossi$paro == "yes"] <- 1
rossi$paro[rossi$paro == "no"] <- 0
rossi$paro <- as.numeric(rossi$paro)
rossi$educ <- as.factor(rossi$educ)
rossi$age[rossi$age <=20 ]<-0
rossi$age[20<rossi$age & rossi$age<= 27]<-1
rossi$age[rossi$age>27]<-2
rossi$prio[rossi$prio < 3]<-0
rossi$prio[rossi$prio >=3 ] <-1
rossi

rossi$prio<-as.factor(rossi$prio)
rossi$age <- as.factor(rossi$age)
rossi$race <- as.factor(rossi$race)
rossi$wexp <- as.factor(rossi$wexp)
rossi$mar <- as.factor(rossi$mar)
```


Document introduction :

A data frame with 432 observations on the following 62 variables. 

week
week of first arrest after release or censoring; all censored observations are censored at 52 weeks.

arrest
1 if arrested, 0 if not arrested.

fin
financial aid: no - 0  yes - 1.

age
in years at time of release.

race
black - 1  other - 0.

wexp
full-time work experience before incarceration: no - 0 or yes - 1.

mar
marital status at time of release: married - 1  not married - 0 .

paro
released on parole? no - 0 or yes - 1.

prio
number of convictions prior to current incarceration.

educ
level of education: 2 = 6th grade or less; 3 = 7th to 9th grade; 4 = 10th to 11th grade; 5 = 12th grade; 6 = some college.

```{r}
```


```{r}
summary(rossi$prio)
```


First we calculate the KM estimate of the survival function.

```{r}
rossi.surv <- Surv(rossi$week,rossi$arrest)
rossi.fit <- survfit(rossi.surv~1)

plot(rossi.fit,mark=19,lwd=2, 
     col=c("steelblue","pink"),
     ylab="Survival Probability", 
     xlab="Days until re-arrested", 
     main="Kaplan-Meier Estimate for rossi data")

```

Compare the Kaplan-Meier estimates of the survivor functions for each covariate individually.

```{r}

kmfin <- survfit(rossi.surv~rossi$fin)
plot(kmfin,mark=19,
     ylab = "Survival Probability",xlab = "survival Time in weeks",
     main = "KM curves for receiving financial aid",col = c(3,6))

legend(0,0.18,c("None","Prsence"),fill=c(3,6))
```


```{r}
kmage <- survfit(rossi.surv~rossi$age)
plot(kmage,mark=19,
     ylab = "Survival Probability",xlab = "survival Time in weeks",
     main = "KM curves for release age",col = c(3,4,5))
legend(0,0.4,c("age <= 20","age between 20 and 27","age>27"),fill=c(3,4,5))

```

```{r}
kmrace <- survfit(rossi.surv~rossi$race)
plot(kmrace,mark=19,
     ylab = "Survival Probability",xlab = "survival Time in weeks",
     main = "KM curves for race",col = c(3,6))
legend(0,0.18,c("other","black"),fill=c(3,6))
```

```{r}
kmwexp <- survfit(rossi.surv~rossi$wexp)
plot(kmwexp,mark=19,
     ylab = "Survival Probability",xlab = "survival Time in weeks",
     main = "KM curves for full time working experience",col = c(3,6))
legend(0,0.2,c("no full time work experience","full time work experience"),fill=c(3,6))
```

```{r}
kmar <- survfit(rossi.surv~rossi$mar)
plot(kmar,mark=19,
     ylab = "Survival Probability",xlab = "survival Time in weeks",
     main = "KM curves for marriage",col = c(3,6))
legend(0,0.2,c("not married","married"),fill=c(3,6))
```

```{r}
kmparo <- survfit(rossi.surv~rossi$paro)
plot(kmparo,mark=19,
     ylab = "Survival Probability",xlab = "survival Time in weeks",
     main = "KM curves for released on paro",col = c(3,6))
legend(0,0.2,c("not released on parole","released on parole"),fill=c(3,6))
```

```{r}
kmprio <- survfit(rossi.surv~rossi$prio)
plot(kmprio,mark=19,
     ylab = "Survival Probability",xlab = "survival Time in weeks",
     main = "KM curves for prior to current incarceration",col = c(3,6))
legend(0,0.2,c("less than 3 conviction","greater or equal to 3 convictions"),fill=c(3,6))
```



```{r}
kmedu <- survfit(rossi.surv~rossi$educ)
plot(kmedu,mark=19,lwd=3,
     ylab = "Survival Probability",xlab = "survival Time in weeks",
     main = "KM curves for education leves",col = c(1,2,3,4,5))
legend(0,0.4,c("6th grade or less","7th to 9th grade","10th to 11th grade","12th grade","some college"),fill = c(1,2,3,4,5))
```


Log-rank test for each covariate and interpretation.

```{r}
fin.diff <- survdiff(rossi.surv~rossi$fin)
fin.diff
```


```{r}
age.diff <- survdiff(rossi.surv~rossi$age)
age.diff
```


```{r}
race.diff <- survdiff(rossi.surv~rossi$race)
race.diff
```


```{r}
wexp.diff <- survdiff(rossi.surv~rossi$wexp)
wexp.diff

```


```{r}
mar.diff <- survdiff(rossi.surv~rossi$mar)
mar.diff

```
```{r}
prio.diff <- survdiff(rossi.surv~rossi$prio)
prio.diff


```



```{r}
paro.diff <- survdiff(rossi.surv~rossi$paro)
paro.diff

```

```{r}
educ.diff <- survdiff(rossi.surv~rossi$educ)
educ.diff
```

Model selection



```{r}
model1 <- coxph(Surv(week,arrest)~fin+age+prio,data=rossi)
cox.zph(model1)
```

From the coxzph test, we want to stratify age variable.

C-log-log plot

```{r}
# financial aid
plot(survfit(Surv(week,arrest)~fin,data=rossi),lwd=2,col=c(2,4),fun="cloglog")
legend(10,1,c("no","yes"),fill=c(2,4))
```

```{r}
plot(survfit(Surv(week,arrest)~age,data=rossi),lwd=2,col=c(2,3,4,5),fun="cloglog")
legend(1,-1,c("less than 20","between 20 and 24","between 24 and 27","greater than 27"),fill=c(2,3,4,5))
```



```{r}
plot(survfit(Surv(week,arrest)~prio,data=rossi),lwd=2,col=c(2,4),fun="cloglog")

```


```{r}
finalmodel <- coxph(Surv(week,arrest)~fin+prio+strata(age),data=rossi)
summary(finalmodel)
```

Interaction term

```{r}
interaction1 <- coxph(Surv(week,arrest)~fin*prio,data=rossi)
anova(interaction1)

```

```{r}
interaction2 <- coxph(Surv(week,arrest)~fin*strata(age),data=rossi)
anova(interaction2)
```

```{r}
interaction3 <- coxph(Surv(week,arrest)~prio*strata(age),data=rossi)
anova(interaction3)
```


```{r}
library(survminer)
library(dplyr)
require("survival")
fit3 <- coxph(Surv(week,arrest)~fin+prio,data=rossi)
ci <- ggforest(fit3,data=rossi)
ci
```
```{r}
finalmodel2 <- coxph(Surv(week,arrest)~fin+prio+strata(age)+strata(age)*prio,data=rossi)
summary(finalmodel2)
```


```{r}
plot(survfit(finalmodel),fun="cumhaz",lwd=5,mark=5,col=c(3,4,5))
legend(0,0.3,c("age less than 20","age between 20 and 27","age greater than 27"),fill = c(3,4,5))
```


AFT model

```{r}
fit_aft <- survreg(Surv(week,arrest)~fin+prio+age,data=rossi)
fitted_values <- fit_aft$linear.predictors
resids <- (log(fit_aft$y[,1]-fitted_values)/fit_Aft$scale)
xx<-seq(min(resids),max(resids),length.out=100)
yy<-plogis(xx, lower.tail = FALSE)

           
```


```{r}
surv.fit=survfit(rossi.surv~fin,data=rossi)
plot(log(surv.fit$time),log(surv.fit$surv/(1-surv.fit$surv)),  type="p", xlab="Log Time", ylab="Log Odds of Survival",col=4,main="Survival odds plot",pch=19,ylim=c(0,15)) 
```


```{r}
surv.fit=survfit(rossi.surv~age,data=rossi)
plot(log(surv.fit$time),log(surv.fit$surv/(1-surv.fit$surv)),  type="p", xlab="Log Time", ylab="Log Odds of Survival",col=3,main="Survival odds plot",pch=19,ylim=c(0,15)) 

```



```{r}
surv.fit=survfit(rossi.surv~prio,data=rossi)
plot(log(surv.fit$time),log(surv.fit$surv/(1-surv.fit$surv)),  type="p", xlab="Log Time", ylab="Log Odds of Survival",col=2,main="Survival odds plot",pch=19,ylim=c(0,15)) 

```

```{r}


```



```{r}
logfinalmodel <- survreg(coxph(Surv(week,arrest)~fin+prio+age,data=rossi),data=rossi,dist = "loglogistic")
summary(logfinalmodel)

```



```{r}

```

